---
title: "Covid-19 Cooperation"
output:
  html_document:
    self_contained: false
    keep_md: true
    code_folding: hide
  pdf_document: default
---

## Import library
```{r, results="hide", warning=FALSE, message=FALSE, include=F}
packages <- c("dplyr", "tibble", "tidyr", "purrr", "FactoMineR", "ggplot2", "lm.beta", "olsrr", "caret", "tidyverse", "reticulate", "factoextra", "MASS", "ggeffects", "effects", "rio", "foreign", "Hmisc","reshape2", "lavaan")

installed_packages <- packages %in% row.names(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed.packages])
}

lapply(packages, library, character.only = TRUE)

``` 

### Read data
```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
setwd("~/2021/Artigo_Covid-19_Daniela")
# import worksheets
data_list <- import_list("Covid-19_Database_Final_backup.xlsx")
```

### Clean tarefa principal
```{r, results="hide", warning=FALSE, message=FALSE, echo=FALSE}

# tarefa principal
tarefa_principal <- data_list$`Tarefa Principal`
glimpse(tarefa_principal)

# rename dv
names(tarefa_principal)[which(names(tarefa_principal) == "Indique, tendo em conta a situação apresentada se iria trabalhar ou não ?")] <- "dv"

# time pressure as factor
tarefa_principal$`Time pressure`<- factor(tarefa_principal$`Time pressure`, levels = c("0","1"))

# Scenario recode
tarefa_principal$Scnario[tarefa_principal$Scnario==1]<-"a"
tarefa_principal$Scnario[tarefa_principal$Scnario==2]<-"b"
tarefa_principal$Scnario[tarefa_principal$Scnario==3]<-"c"
tarefa_principal$Scnario[tarefa_principal$Scnario==4]<-"d"

# Scenario as factor
tarefa_principal$Scnario <- factor(tarefa_principal$Scnario, levels = c("a","b","c","d"))

# na.removal
tarefa_principal<- tarefa_principal[!is.na(tarefa_principal$dv), ]
tarefa_principal<- tarefa_principal[!is.na(tarefa_principal$`Time pressure`),]
tarefa_principal<- tarefa_principal[!is.na(tarefa_principal$Scnario),]
```

### clean risk/probability discount & merge
```{r, results="hide", warning=FALSE, message=FALSE}
# risk/probability discount 
data_list$Discount$R2 <- as.numeric(as.character(data_list$Discount$R2))

risk_tarefa <- data_list$`Discount`
tarefa_principal$AUC = risk_tarefa[match(tarefa_principal$Code, risk_tarefa$Code),"AUC"]

# rem outliers
yout = tarefa_principal$AUC < 0.07

#boxplot(tarefa_principal$AUC,
  #ylab = "AUC"
#)
```

### COVID_Risk PCA 
```{r, warning=FALSE, message=FALSE}

# covid risk questionnaire
covid<- data_list$`Covid Risk questionário`
covid<-na.omit(covid)

# rename col
colnames(covid)[2]<-"v1"
colnames(covid)[3]<-"v2"
colnames(covid)[4]<-"v3"
colnames(covid)[5]<-"v4"
colnames(covid)[6]<-"v5"
colnames(covid)[7]<-"v6"
colnames(covid)[8]<-"v7"
colnames(covid)[9]<-"v8"
colnames(covid)[10]<-"v9"
colnames(covid)[11]<-"v10"
colnames(covid)[12]<-"v11"
#names(covid)

#pca
pca <- PCA(covid[,3:12], graph=F)
summary(pca,nbelements = 10, ncp = 3)
get_eig(pca)

fviz_screeplot(pca, addlabels = TRUE, ylim = c(0, 50))
var <- get_pca_var(pca)
head(var)
fviz_pca_var(pca, col.var="contrib",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE 
)
```

### CFA COVID-19 scale 
```{r, warning=FALSE, message=FALSE}
# cfa
model<-'
c1=~v2+v4+v9
c2=~v3+v7+v8+v11
c3=~v10+v5'
fit<- cfa(model, data=covid)
summary(fit, fit.measures=T,standardized=T)
```

### Merge COVID to tarefa_principal
```{r, warning=FALSE, message=FALSE}
# dimensions 

covid$spread <- rowMeans(covid[,c("v2", "v4","v9")], na.rm=TRUE)

covid$impact <- rowMeans(covid[,c("v3", "v7", "v8", "v11")], na.rm=TRUE)

covid$distant_spread <- rowMeans(covid[,c("v10", "v5")], na.rm=TRUE)

covid$impact_to_other <- covid$v1

# merge to tarefa principal
names(covid)[1]<-"Code"

tarefa_principal$i1=covid[match(tarefa_principal$Code, covid$Code),"v1"]
tarefa_principal$i2=covid[match(tarefa_principal$Code, covid$Code),"v2"]
tarefa_principal$i3=covid[match(tarefa_principal$Code, covid$Code),"v3"]
tarefa_principal$i4=covid[match(tarefa_principal$Code, covid$Code),"v4"]
tarefa_principal$i5=covid[match(tarefa_principal$Code, covid$Code),"v5"]
tarefa_principal$i6=covid[match(tarefa_principal$Code, covid$Code),"v6"]
tarefa_principal$i7=covid[match(tarefa_principal$Code, covid$Code),"v7"]
tarefa_principal$i8=covid[match(tarefa_principal$Code, covid$Code),"v8"]
tarefa_principal$i9=covid[match(tarefa_principal$Code, covid$Code),"v9"]
tarefa_principal$i10=covid[match(tarefa_principal$Code, covid$Code),"v10"]
tarefa_principal$i11=covid[match(tarefa_principal$Code, covid$Code),"v11"]
tarefa_principal$i12=covid[match(tarefa_principal$Code, covid$Code),"v12"]


tarefa_principal$covid_spread = covid[match(tarefa_principal$Code, covid$Code),"spread"]
tarefa_principal$covid_impact = covid[match(tarefa_principal$Code, covid$Code), "impact"]
tarefa_principal$covid_distant_spread = covid[match(tarefa_principal$Code, covid$Code), "distant_spread"]
tarefa_principal$covid_impact_to_other = covid[match(tarefa_principal$Code, covid$Code), "impact_to_other"]
```

### IRI clean & merge
```{r, results="hide", warning=FALSE, message=FALSE}
# IRI scale
IRI<- data_list$`IRI tarefa`
#str(IRI)
names(IRI)[1]<-"Code"
#names(IRI)

# merge to tarefa principal 
tarefa_principal$iri_TP = IRI[match(tarefa_principal$Code, IRI$Code),"TP (Tomada de Perspectiva)"]
tarefa_principal$iri_PE = IRI[match(tarefa_principal$Code, IRI$Code),"Preocupação Empatica (PE)"]
tarefa_principal$iri_DP = IRI[match(tarefa_principal$Code, IRI$Code),"Desconforto Pessoal (DP)"]
tarefa_principal$iri_F = IRI[match(tarefa_principal$Code, IRI$Code),"Fantasia (F)"]
tarefa_principal$iri_E = IRI[match(tarefa_principal$Code, IRI$Code),"Empatia"]
```

### Preliminary Sociodemographics & merge
```{r, results="hide", warning=FALSE, message=FALSE}
# sociodemographic data 
sociodemog<-data_list$`1a parte_ Sociodemografico`
#names(sociodemog)
names(sociodemog)[1]<-"Code"

# merge to tarefa principal
tarefa_principal$age = sociodemog[match(tarefa_principal$Code, IRI$Code),"Idade:"]
tarefa_principal$gender = sociodemog[match(tarefa_principal$Code, IRI$Code),"Género:"]
tarefa_principal$socioeconomic= sociodemog[match(tarefa_principal$Code, sociodemog$Code), "Qual é o seu nível sócio-económico?"]
tarefa_principal$education = sociodemog[match(tarefa_principal$Code, sociodemog$Code), "Habilitações literárias: - Selected Choice"]
tarefa_principal$job = sociodemog[match(tarefa_principal$Code, sociodemog$Code),"Profissão:"]

# as.factor                                        
tarefa_principal$gender<-factor(tarefa_principal$gender)
tarefa_principal$socioeconomic <- factor(tarefa_principal$socioeconomic)
tarefa_principal$education <- factor(tarefa_principal$education)
tarefa_principal$job <- factor(tarefa_principal$job)
tarefa_principal$fase<-factor(tarefa_principal$`Fase de resposta`)
```

### Exploratory plots

## Interaction
```{r,dev=c('png', 'tiff'), warning=FALSE, message=FALSE}
xlabs <- c("Control", "Impact \n  Uncertainty", "Worst-Case", "Indirect \n  Transmission")

ggplot(data=tarefa_principal,                     
       aes(x=Scnario, 
           y=dv, 
           fill=`Time pressure`)) +                       
  geom_boxplot() +                            
  labs(y='Likelihood Of Going To Work') + xlab("Scenario") +                
  theme_bw() + 
    scale_x_discrete(labels= xlabs) +   
  theme(axis.text.x = element_text(color = "black")) +
  #removing legend title and renaming labels
  scale_fill_discrete(name="",
         labels=c("Without Time Limit","Time Pressure")) + 
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```

### Scenario
```{r,dev=c('png', 'tiff'), warning=FALSE, message=FALSE}
#create x labels
xlabs <- c("Control", "Impact Uncertainty", "Worst-Case","Indirect Transmission")

ggplot(data=tarefa_principal,                     
       aes(x=Scnario, 
           y=dv)) +                       
  geom_boxplot(fill="seashell2") +                      
  labs(y='Likelihood Of Going To Work') + xlab("Scenario") +
   theme(axis.text.x = element_text(color = "black")) +
    theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  scale_x_discrete(labels= xlabs) +  
  theme(axis.text.x = element_text(color = "black"))
```

### Time pressure
```{r,dev=c('png', 'tiff'), warning=FALSE, message=FALSE}
xlabs<-c("Without Time Limit","Time Pressure")

ggplot(data=tarefa_principal,                     
       aes(x=`Time pressure`, 
           y=dv)) +                       
  geom_boxplot(fill = "SKYblue") +
  labs(y='Likelihood Of Going To Work') +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
    scale_x_discrete(labels= xlabs) +
  theme(axis.text.x = element_text(color = "black")) + 
  theme_bw() + 
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
  
```

## Statistical Analysis 

### preprocessing
```{r, warning=FALSE, message=FALSE}
tarefa_principal$dv_factor<-factor(tarefa_principal$dv, ordered = T)
#colnames(tarefa_principal)
tarefa_principal_2<-tarefa_principal[, c(1,2,3,18,25,37,38,39,40,41,42,43,44,45,46,47,48,49,50,51,52)]
tarefa_principal_2_na.rem<- na.omit(tarefa_principal_2)

#collapse education levels, since there were only 3 participants below 9th grade afert na's removal

#highschool_or_below
tarefa_principal_2_na.rem$education[tarefa_principal_2_na.rem$education==2]<-"2"
tarefa_principal_2_na.rem$education[tarefa_principal_2_na.rem$education==3]<-"2"
tarefa_principal_2_na.rem$education[tarefa_principal_2_na.rem$education==4]<-"2"

#bachelor_or_higher
tarefa_principal_2_na.rem$education[tarefa_principal_2_na.rem$education==5]<-"3"
tarefa_principal_2_na.rem$education[tarefa_principal_2_na.rem$education==6]<-"3"
tarefa_principal_2_na.rem$education[tarefa_principal_2_na.rem$education==7]<-"3"

#collapse low and moderate/low socioeconomic levels, since there were only 5 participants with low socioeconomic levels

tarefa_principal_2_na.rem$socioeconomic[tarefa_principal_2_na.rem$socioeconomic==1]<-"1"
tarefa_principal_2_na.rem$socioeconomic[tarefa_principal_2_na.rem$socioeconomic==2]<-"1"
tarefa_principal_2_na.rem$socioeconomic[tarefa_principal_2_na.rem$socioeconomic==3]<-"2"
tarefa_principal_2_na.rem$socioeconomic[tarefa_principal_2_na.rem$socioeconomic==4]<-"3"

tarefa_principal_2_na.rem$job[tarefa_principal_2_na.rem$job==1]<-"1"
tarefa_principal_2_na.rem$job[tarefa_principal_2_na.rem$job==2]<-"2"
tarefa_principal_2_na.rem$job[tarefa_principal_2_na.rem$job==3]<-"3"
tarefa_principal_2_na.rem$job[tarefa_principal_2_na.rem$job==4]<-"4"
tarefa_principal_2_na.rem$job[tarefa_principal_2_na.rem$job==6]<-"6"
```

### fit model (dv ~ Time pressure*Scnario + fase)
```{r, warning=FALSE, message=FALSE}
data_olr <- tarefa_principal[, c("Time pressure", "Scnario", "Fase de resposta", "dv_factor")]
data_olr$fase <- data_olr$`Fase de resposta`

m10<- polr(dv_factor ~`Time pressure` + Scnario + fase, data= data_olr, Hess=TRUE)
summary(m10)

ctable <- coef(summary(m10))
p <- pnorm(abs(ctable[, "t value"]), lower.tail = F) * 2
ctable <- cbind(ctable, "p value" = p)
ctable
ci <- confint(m10)
exp(cbind(OR = coef(m10), ci))
```

#subset participants 
```{r}
#filter
tarefa_principal = subset(tarefa_principal, !(Code %in% c("13051995CB","01061995RR","0860","8001AD","8232ai","1757AS","5485mj","1876SD","9228+IS","5986","5609 M. F.","7687MM","1682BP","579","8260JM","José","3557DM","2516BS","6309 A M","2130MJ","6336RA","2686","3586AI","0243IC","9878JI","28091998IS","01091993JM","05112001EP","0601pf","1076MI","4278RM","4744CS","1336SF","2861AC","7505DP","9044jr","3560HM","7784","9087AR","7631AR","5371pm","0168 cd","7771MG","6872mm","3514 + MS","7779BA","1010DE","6760AP")))

data_olr <- tarefa_principal[, c("Time pressure", "Scnario", "Fase de resposta", "dv_factor")]
data_olr$fase <- data_olr$`Fase de resposta`

m10<- polr(dv_factor ~`Time pressure` + Scnario + fase, data= data_olr, Hess=TRUE)
summary(m10)

ctable <- coef(summary(m10))
p <- pnorm(abs(ctable[, "t value"]), lower.tail = F) * 2
ctable <- cbind(ctable, "p value" = p)
ctable
ci <- confint(m10)
exp(cbind(OR = coef(m10), ci))
```


### random sampling
```{r}
temp = data_olr
#data_olr$`Time pressure` = as.character(data_olr$`Time pressure`)
df_0 = subset(data_olr, `Time pressure` == 0)
df_1 = subset(temp, `Time pressure` == 1)
set.seed(12345)    
df_0 = sample_n(df_0, 182)
df_random0 = rbind(df_0,df_1)
#df_random0$`Time pressure`=factor(df_random1$`Time pressure`)

m10<- polr(dv_factor ~`Time pressure` + Scnario + fase, data= df_random0, Hess=TRUE)
summary(m10)

ctable <- coef(summary(m10))
p <- pnorm(abs(ctable[, "t value"]), lower.tail = F) * 2
ctable <- cbind(ctable, "p value" = p)
ctable
ci <- confint(m10)
exp(cbind(OR = coef(m10), ci))
```

```{r}
temp = data_olr
#data_olr$`Time pressure` = as.character(data_olr$`Time pressure`)
df_0 = subset(data_olr, `Time pressure` == 0)
df_1 = subset(temp, `Time pressure` == 1)
set.seed(54321)    
df_0 = sample_n(df_0, 182)
df_random0 = rbind(df_0,df_1)
#df_random0$`Time pressure`=factor(df_random1$`Time pressure`)

m10<- polr(dv_factor ~`Time pressure` + Scnario + fase, data= df_random0, Hess=TRUE)
summary(m10)

ctable <- coef(summary(m10))
p <- pnorm(abs(ctable[, "t value"]), lower.tail = F) * 2
ctable <- cbind(ctable, "p value" = p)
ctable
ci <- confint(m10)
exp(cbind(OR = coef(m10), ci))
```

```{r}
temp = data_olr
#data_olr$`Time pressure` = as.character(data_olr$`Time pressure`)
df_0 = subset(data_olr, `Time pressure` == 0)
df_1 = subset(temp, `Time pressure` == 1)
set.seed(1000)    
df_0 = sample_n(df_0, 182)
df_random0 = rbind(df_0,df_1)
#df_random0$`Time pressure`=factor(df_random1$`Time pressure`)

m10<- polr(dv_factor ~`Time pressure` + Scnario + fase, data= df_random0, Hess=TRUE)
summary(m10)

ctable <- coef(summary(m10))
p <- pnorm(abs(ctable[, "t value"]), lower.tail = F) * 2
ctable <- cbind(ctable, "p value" = p)
ctable
ci <- confint(m10)
exp(cbind(OR = coef(m10), ci))
```

### Ordinal Logistic Regression (other variables) 254 participants
```{r, warning=FALSE, message=FALSE}
m <- polr(dv_factor ~  AUC + covid_distant_spread + covid_impact_to_other + covid_spread + covid_impact + iri_TP + iri_PE + iri_DP + iri_F + iri_E + age + gender + socioeconomic, data= tarefa_principal_2_na.rem, Hess=TRUE)
summary(m)
ctable <- coef(summary(m))
p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
ctable <- cbind(ctable, "p value" = p)
ctable
#ci <- confint(m)
exp(cbind("Odds ratio" = coef(m), confint.default(m, level = 0.95)))
#exp(coef(m))

```

### kruskal-wallis group
```{r message=FALSE, warning=FALSE}

#8 conditions (Scenario*Time pressure)
dataset <- mutate(tarefa_principal_2_na.rem,
        newvar = case_when(
                Scnario ==  "a" & `Time pressure`== "0" ~  1, 
                Scnario ==  "b" & `Time pressure`== "0" ~  2,
                Scnario ==  "c" & `Time pressure`== "0" ~  3,
                Scnario ==  "d" & `Time pressure`== "0" ~  4,
                Scnario ==  "a" & `Time pressure`== "1" ~  5,
                Scnario ==  "b" & `Time pressure`== "1" ~  6,
                Scnario ==  "c" & `Time pressure`== "1" ~  7,
                Scnario ==  "d" & `Time pressure`== "1" ~  8,
               TRUE ~ NA_real_ # This is for all other values 
             ))                # not covered by the above.

dataset$newvar<- factor(dataset$newvar)

#see if there are differences for each variable between experimental conditions
fit <- lapply(dataset[,c("AUC","covid_distant_spread", "covid_impact_to_other", "covid_spread", "covid_impact", "iri_TP", "iri_PE", "iri_DP", "iri_F", "iri_E", "age", "gender", "education", "socioeconomic")], function(y) kruskal.test(y ~ newvar, data = dataset))
fit

#post-hoc covid_spread
fit2<- dunnTest(dataset$covid_spread~dataset$newvar, method = dunn.test::p.adjustment.methods[c(4,
  2:3, 5:8, 1)])
fit2 

#post-hoc socioeconomic
dataset$socioeconomic<- as.numeric(dataset$socioeconomic)
fit2<- dunnTest(dataset$socioeconomic~dataset$newvar, method = dunn.test::p.adjustment.methods[c(4,
  2:3, 5:8, 1)])
fit2 
```

#OLR Plot
```{python, include=F}
df = r.tarefa_principal
df.dtypes
```

```{python}
import pandas as pd 
import seaborn as sns 
import matplotlib as plt
from sklearn.ensemble import RandomForestClassifier
from sklearn.svm import SVC
from sklearn import svm
from sklearn.neural_network import MLPClassifier
from sklearn.metrics import confusion_matrix, classification_report
from sklearn.preprocessing import StandardScaler, LabelEncoder
from sklearn.model_selection import train_test_split
from sklearn.preprocessing import scale 
from sklearn import decomposition
from matplotlib import pyplot as plt
import plotly.express as px
import numpy as np
from sklearn.model_selection import cross_val_score, StratifiedKFold, KFold
from sklearn.linear_model import LogisticRegression
from sklearn.metrics import accuracy_score
from sklearn.model_selection import GridSearchCV
import statsmodels.formula.api as smf
import textwrap
import scipy.stats as stats
```

```{python}
predict, validate = train_test_split(df, test_size=0.2, random_state=42)
```
```{python}
mod = smf.ols(formula='dv ~ i2 + i3 + i4 + i5 + i6 + i7 + i8 + i9 + i10+i11', data=df)

fit = mod.fit()
fit.summary()

```

```{python}
mod_discovery = smf.ols(formula='dv ~ i2 + i3 + i4 + i5 + i6 + i7 + i8 + i9 + i10 + i11', data=predict)

res_discovery = mod_discovery.fit()
res_discovery.summary()

```

```{python, include = F}
mod_validation = smf.ols(formula='dv ~ i2 + i3 + i4 + i5 + i6 + i7 + i8 + i9 + i10 + i11', data=validate)

res_validation = mod_validation.fit()
```
```{python}
names = ['How likely do you think you are to catch the virus','How badly do you think you will be affected economically if you specifically catch the virus? (for example through loss of work)','How likely do you think it is that a loved one will become infected?','How likely do you think the average person in your state is to become infected?','If you do contract the virus, how likely do you think it is that you will pass it on to someone else','How badly do you think your health will be affected if you do catch the virus?','How badly do you think you will be affected by the global effects of the virus (for example economic recession, reduced healthcare capacity)?','How likely do you think the average person in your neighbourhood is to become infected?','How likely do you think the average person in your country is to become infected?','If you do contract the virus and pass it on to someone else, how badly do you think they would be affected?']


# font type & size
plt.rcParams.update({'font.size': 12})
plt.rcParams["font.family"] = "Times New Roman"

fig, ax = plt.subplots(figsize=(10,14))
fig.set_tight_layout(True)

# DISCOVERY
ax.axvline(0, linestyle=':', color='gray', linewidth=2)

# Points and error bars
ax.errorbar(y=np.arange(len(res_discovery.params[1:])) - 0.1, x=res_discovery.params[1:], 
             xerr=np.abs(res_discovery.conf_int().values[1:, :] - res_discovery.params[1:, np.newaxis]).T, 
             color='black', linewidth=2, fmt='^', label='estimation set')

ax.errorbar(y=np.arange(len(res_validation.params[1:])) + 0.1, x=res_validation.params[1:], 
             xerr=np.abs(res_validation.conf_int().values[1:, :] - res_validation.params[1:, np.newaxis]).T, 
             color='gray', linewidth=2, fmt='s', label='validation set')

# Labels
ax.set_yticks(range(len(names)))
ax.set_yticklabels(['\n'.join(textwrap.wrap(q, 60, break_long_words=False)) for q in names])

## Titles 
ax.set_xlabel('Regression coefficient (+/- 95% CI)')
ax.legend(loc=7)

plt.show()

##save
from PIL import Image
from io import BytesIO
# (1) save the image in memory in PNG format
png1 = BytesIO()
fig.savefig(png1, format='png')

# (2) load this image into PIL
png2 = Image.open(png1)

# (3) save as TIFF
png2.save('glm_Daniela_graph3.tiff')
png1.close()
```
