---
title: "A dual process approach to prosocial behaviour"
author: "Nuno Fernandes"
date: "23/11/2021"
output:
  rmarkdown::html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Import library
```{r, results="hide", warning=FALSE, message=FALSE}
packages <- c("dplyr", "tibble", "tidyr", "purrr", "FactoMineR", "ggplot2", "lm.beta", "olsrr", "rcompanion", "FSA", "caret", "tidyverse", "reticulate", "factoextra", "MASS", "ggeffects", "effects", "rio", "foreign", "Hmisc","reshape2", "misty", "lavaan")

installed_packages <- packages %in% row.names(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed.packages])
}

lapply(packages, library, character.only = TRUE)

``` 

### Read data
```{r, results='hide', warning=FALSE, message=FALSE}
setwd("~/2021/Artigo_Covid-19_Daniela")
# import worksheets
data_list <- import_list("df_cooperation_covid.xlsx")
```

### Clean tarefa principal
```{r, results="hide", warning=FALSE, message=FALSE}

# tarefa principal
main_task <- data_list$task
glimpse(main_task)

# rename dv
names(main_task)[which(names(main_task) == "Prosocial behavior")] <- "dv"

# `Time pressure` as factor
main_task$`Time pressure`<- factor(main_task$`Time pressure`, levels = c("0","1"))

# Scenario recode
main_task$Scenario[main_task$Scenario==1]<-"a"
main_task$Scenario[main_task$Scenario==2]<-"b"
main_task$Scenario[main_task$Scenario==3]<-"c"
main_task$Scenario[main_task$Scenario==4]<-"d"

# Scenario as factor
main_task$Scenario <- factor(main_task$Scenario, levels = c("a","b","c","d"))

# na.removal
main_task<- main_task[!is.na(main_task$dv), ]
main_task<- main_task[!is.na(main_task$`Time pressure`),]
main_task<- main_task[!is.na(main_task$Scenario),]
```

### Clean risk/probability discount & merge
```{r, results="hide", warning=FALSE, message=FALSE}
# risk/probability discount 
data_list$Discount$R2 <- as.numeric(as.character(data_list$Discount$R2))

risk_tarefa <- data_list$`Discount`
main_task$AUC = risk_tarefa[match(main_task$Code, risk_tarefa$Code),"AUC"]

# rem outliers
yout = main_task$AUC < 0.07

#boxplot(main_task$AUC,
  #ylab = "AUC"
#)
```

### Clean SVO and merge
```{r}
svo <- data_list$`svo task`
main_task$SVO = svo[match(main_task$Code, risk_tarefa$Code), "SVO"]
main_task$SVO <- as.factor(main_task$SVO)
```


### COVID_Risk PCA 
```{r, warning=FALSE, message=FALSE}

# covid risk questionnaire
covid<- data_list$` Subj Risk COVID-19`
covid<-na.omit(covid)

# rename col
colnames(covid)[2]<-"v1"
colnames(covid)[3]<-"v2"
colnames(covid)[4]<-"v3"
colnames(covid)[5]<-"v4"
colnames(covid)[6]<-"v5"
colnames(covid)[7]<-"v6"
colnames(covid)[8]<-"v7"
colnames(covid)[9]<-"v8"
colnames(covid)[10]<-"v9"
colnames(covid)[11]<-"v10"
colnames(covid)[12]<-"v11"
#names(covid)

#pca
pca <- PCA(covid[,3:12], graph=F)
summary(pca,nbelements = 10, ncp = 3)
get_eig(pca)

fviz_screeplot(pca, addlabels = TRUE, ylim = c(0, 50))
var <- get_pca_var(pca)
head(var)
fviz_pca_var(pca, col.var="contrib",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE 
)
```

### CFA COVID-19 scale 
```{r, warning=FALSE, message=FALSE}
# cfa
model<-'
c1=~v2+v4+v9
c2=~v3+v7+v8+v11
c3=~v10+v5'
fit<- cfa(model, data=covid)
summary(fit, fit.measures=T,standardized=T)
```

### Merge COVID to main_task
```{r, warning=FALSE, message=FALSE}
# dimensions 

covid$spread <- rowMeans(covid[,c("v2", "v4","v9")], na.rm=TRUE)

covid$impact <- rowMeans(covid[,c("v3", "v7", "v8", "v11")], na.rm=TRUE)

covid$distant_spread <- rowMeans(covid[,c("v10", "v5")], na.rm=TRUE)

covid$impact_to_other <- covid$v1

# merge to tarefa principal
names(covid)[1]<-"Code"

main_task$i1=covid[match(main_task$Code, covid$Code),"v1"]
main_task$i2=covid[match(main_task$Code, covid$Code),"v2"]
main_task$i3=covid[match(main_task$Code, covid$Code),"v3"]
main_task$i4=covid[match(main_task$Code, covid$Code),"v4"]
main_task$i5=covid[match(main_task$Code, covid$Code),"v5"]
main_task$i6=covid[match(main_task$Code, covid$Code),"v6"]
main_task$i7=covid[match(main_task$Code, covid$Code),"v7"]
main_task$i8=covid[match(main_task$Code, covid$Code),"v8"]
main_task$i9=covid[match(main_task$Code, covid$Code),"v9"]
main_task$i10=covid[match(main_task$Code, covid$Code),"v10"]
main_task$i11=covid[match(main_task$Code, covid$Code),"v11"]
main_task$i12=covid[match(main_task$Code, covid$Code),"v12"]


main_task$covid_spread = covid[match(main_task$Code, covid$Code),"spread"]
main_task$covid_impact = covid[match(main_task$Code, covid$Code), "impact"]
main_task$covid_distant_spread = covid[match(main_task$Code, covid$Code), "distant_spread"]
main_task$covid_impact_to_other = covid[match(main_task$Code, covid$Code), "impact_to_other"]
```

### IRI clean & merge
```{r, results="hide", warning=FALSE, message=FALSE}
# IRI scale
IRI<- data_list$`IRI task`
#str(IRI)
names(IRI)[1]<-"Code"
#names(IRI)

# merge to tarefa principal 
main_task$iri_TP = IRI[match(main_task$Code, IRI$Code),"PT"]
main_task$iri_PE = IRI[match(main_task$Code, IRI$Code),"EC"]
main_task$iri_DP = IRI[match(main_task$Code, IRI$Code),"PD"]
main_task$iri_F = IRI[match(main_task$Code, IRI$Code),"F"]
#main_task$iri_E = IRI[match(main_task$Code, IRI$Code),"Empathy"]
```

### Preliminary Sociodemographics & merge
```{r, results="hide", warning=FALSE, message=FALSE}
# sociodemographic data 
sociodemog<-data_list$`Socio-demographic `
#names(sociodemog)
names(sociodemog)[1]<-"Code"

# merge to tarefa principal
main_task$age = sociodemog[match(main_task$Code, IRI$Code),"Age"]
main_task$gender = sociodemog[match(main_task$Code, IRI$Code),"Gender"]
main_task$socioeconomic= sociodemog[match(main_task$Code, sociodemog$Code), "SES"]
main_task$education = sociodemog[match(main_task$Code, sociodemog$Code), "Educational Qualifications"]
main_task$job = sociodemog[match(main_task$Code, sociodemog$Code),"Profession"]

# as.factor                                        
main_task$gender<-factor(main_task$gender)
main_task$socioeconomic <- factor(main_task$socioeconomic)
main_task$education <- factor(main_task$education)
main_task$job <- factor(main_task$job)
main_task$phase<-factor(main_task$Phase)
```

### Exploratory plots

### Interaction
```{r,dev=c('png', 'tiff'), warning=FALSE, message=FALSE}
xlabs <- c("Control", "Impact \n  Uncertainty", "Worst-Case", "Indirect \n  Transmission")

ggplot(data=main_task,                     
       aes(x=Scenario, 
           y=dv, 
           fill=`Time pressure`)) +                       
  geom_boxplot() +                            
  labs(y='Likelihood Of Going To Work') + xlab("Scenario") +                
  theme_bw() + 
    scale_x_discrete(labels= xlabs) +   
  theme(axis.text.x = element_text(color = "black")) +
  #removing legend title and renaming labels
  scale_fill_discrete(name="",
         labels=c("Without Time Limit","`Time pressure`")) + 
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```

### Scenario
```{r,dev=c('png', 'tiff'), warning=FALSE, message=FALSE}
#create x labels
xlabs <- c("Control", "Impact Uncertainty", "Worst-Case","Indirect Transmission")

ggplot(data=main_task,                     
       aes(x=Scenario, 
           y=dv)) +                       
  geom_boxplot(fill="seashell2") +                      
  labs(y='Likelihood Of Going To Work') + xlab("Scenario") +
   theme(axis.text.x = element_text(color = "black")) +
    theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  scale_x_discrete(labels= xlabs) +  
  theme(axis.text.x = element_text(color = "black"))
```

### Time pressure
```{r,dev=c('png', 'tiff'), warning=FALSE, message=FALSE}
xlabs<-c("Without Time Limit","`Time pressure`")

ggplot(data=main_task,                     
       aes(x=`Time pressure`, 
           y=dv)) +                       
  geom_boxplot(fill = "SKYblue") +
  labs(y='Likelihood Of Going To Work') +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
    scale_x_discrete(labels= xlabs) +
  theme(axis.text.x = element_text(color = "black")) + 
  theme_bw() + 
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
  
```

## Statistical Analysis 

### preprocessing
```{r, warning=FALSE, message=FALSE}

main_task$dv_factor<-factor(main_task$dv, ordered = T)
#colnames(main_task)
#main_task_2<-main_task[, c(1,2,3,18,25,37,38,39,40,41,42,43,44,45,46,47,48,49,50,51,52)]


main_task_2<-main_task[, c("Code","Time pressure","Scenario","dv","Phase","AUC","i1","i2","i3","i4","i5","i6","i7","i8","i9","i10", "i11", "covid_spread", "covid_impact", "covid_distant_spread","covid_impact_to_other", "iri_TP", "iri_PE", "iri_DP","iri_F","age", "gender", "socioeconomic", "education", "job", "phase", "dv_factor")]
main_task_2_na.rem<- na.omit(main_task_2)

#collapse education levels, since there were only 3 participants below 9th grade afert na's removal

#highschool_or_below
main_task_2_na.rem$education[main_task_2_na.rem$education==2]<-"2"
main_task_2_na.rem$education[main_task_2_na.rem$education==3]<-"2"
main_task_2_na.rem$education[main_task_2_na.rem$education==4]<-"2"

#bachelor_or_higher
main_task_2_na.rem$education[main_task_2_na.rem$education==5]<-"3"
main_task_2_na.rem$education[main_task_2_na.rem$education==6]<-"3"
main_task_2_na.rem$education[main_task_2_na.rem$education==7]<-"3"

#collapse low and moderate/low socioeconomic levels, since there were only 5 participants with low socioeconomic levels

main_task_2_na.rem$socioeconomic[main_task_2_na.rem$socioeconomic==1]<-"1"
main_task_2_na.rem$socioeconomic[main_task_2_na.rem$socioeconomic==2]<-"1"
main_task_2_na.rem$socioeconomic[main_task_2_na.rem$socioeconomic==3]<-"2"
main_task_2_na.rem$socioeconomic[main_task_2_na.rem$socioeconomic==4]<-"3"

main_task_2_na.rem$job[main_task_2_na.rem$job==1]<-"1"
main_task_2_na.rem$job[main_task_2_na.rem$job==2]<-"2"
main_task_2_na.rem$job[main_task_2_na.rem$job==3]<-"3"
main_task_2_na.rem$job[main_task_2_na.rem$job==4]<-"4"
main_task_2_na.rem$job[main_task_2_na.rem$job==6]<-"6"

#SVO for ORDINAL LR
main_task_3 <- main_task[, c("Code","Time pressure","Scenario","dv","Phase","AUC","i1","i2","i3","i4","i5","i6","i7","i8","i9","i10", "i11", "covid_spread", "covid_impact", "covid_distant_spread","covid_impact_to_other", "iri_TP", "iri_PE", "iri_DP","iri_F","age", "gender", "socioeconomic", "education", "job", "phase", "dv_factor", "SVO")]

main_task_3_na.rem<-na.omit(main_task_3)
```

### fit model (dv ~ `Time pressure`*Scenario + phase)
```{r, warning=FALSE, message=FALSE}
data_olr <- main_task[, c("Time pressure", "Scenario", "phase", "dv_factor")]
#data_olr$phase <- data_olr$Phase

m10<- polr(dv_factor ~`Time pressure` + Scenario + phase, data= data_olr, Hess=TRUE)
summary(m10)

ctable <- coef(summary(m10))
p <- pnorm(abs(ctable[, "t value"]), lower.tail = F) * 2
ctable <- cbind(ctable, "p value" = p)
ctable
ci <- confint(m10)
exp(cbind(OR = coef(m10), ci))
```
### Ordinal Logistic Regression (other variables) 254 participants
```{r, warning=FALSE, message=FALSE}
m <- polr(dv_factor ~  AUC + covid_distant_spread + covid_impact_to_other + covid_spread + covid_impact + iri_TP + iri_PE + iri_DP + iri_F + age + gender + socioeconomic, data= main_task_2_na.rem, Hess=TRUE)
summary(m)
ctable <- coef(summary(m))
p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
ctable <- cbind(ctable, "p value" = p)
ctable
#ci <- confint(m)
exp(cbind("Odds ratio" = coef(m), confint.default(m, level = 0.95)))
#exp(coef(m))

```


### kruskal-wallis 
```{r message=FALSE, warning=FALSE}

#8 conditions (Scenario*`Time pressure`)
dataset <- mutate(main_task_2_na.rem,
        newvar = case_when(
                Scenario ==  "a" & `Time pressure`== "0" ~  1, 
                Scenario ==  "b" & `Time pressure`== "0" ~  2,
                Scenario ==  "c" & `Time pressure`== "0" ~  3,
                Scenario ==  "d" & `Time pressure`== "0" ~  4,
                Scenario ==  "a" & `Time pressure`== "1" ~  5,
                Scenario ==  "b" & `Time pressure`== "1" ~  6,
                Scenario ==  "c" & `Time pressure`== "1" ~  7,
                Scenario ==  "d" & `Time pressure`== "1" ~  8,
               TRUE ~ NA_real_ # This is for all other values 
             ))                    # not covered by the above.

dataset$newvar<- factor(dataset$newvar)

#Differences for each variable between experimental conditions


fit <- lapply(dataset[,c("covid_impact_to_other", "gender", "education", "socioeconomic")], function(y) kruskal.test(y ~ newvar, data = dataset))
fit
```
### Kruskal-wallis for SVO
```{r}
dataset2 <- mutate(main_task_3_na.rem,
        newvar = case_when(
                Scenario ==  "a" & `Time pressure`== "0" ~  1, 
                Scenario ==  "b" & `Time pressure`== "0" ~  2,
                Scenario ==  "c" & `Time pressure`== "0" ~  3,
                Scenario ==  "d" & `Time pressure`== "0" ~  4,
                Scenario ==  "a" & `Time pressure`== "1" ~  5,
                Scenario ==  "b" & `Time pressure`== "1" ~  6,
                Scenario ==  "c" & `Time pressure`== "1" ~  7,
                Scenario ==  "d" & `Time pressure`== "1" ~  8,
               TRUE ~ NA_real_ # This is for all other values 
             ))                    # not covered by the above.

kruskal.test(SVO~newvar, data = dataset2)

```
### Dunn post-hoc with bonferroni correction
```{r}
#post-hoc socioeconomic
dataset$socioeconomic<- as.numeric(dataset$socioeconomic)
fit2<- dunnTest(dataset$socioeconomic~dataset$newvar, method = dunn.test::p.adjustment.methods[c(4,
  2:3, 5:8, 1)])
fit2 
  
```

### Anova
```{r}
#Differences for each variable between experimental conditions

fit <- lapply(dataset[,c("age", "AUC","covid_distant_spread", "covid_spread", "covid_impact", "iri_TP", "iri_PE", "iri_DP", "iri_F")], function(y) summary(aov(y ~ Scenario, data = dataset)))

fit
```
### OLR Plot
```{python, include=F}
df = r.main_task
df.dtypes
```

```{python}
import pandas as pd 
import seaborn as sns 
import matplotlib as plt
from sklearn.ensemble import RandomForestClassifier
from sklearn.svm import SVC
from sklearn import svm
from sklearn.neural_network import MLPClassifier
from sklearn.metrics import confusion_matrix, classification_report
from sklearn.preprocessing import StandardScaler, LabelEncoder
from sklearn.model_selection import train_test_split
from sklearn.preprocessing import scale 
from sklearn import decomposition
from matplotlib import pyplot as plt
import plotly.express as px
import numpy as np
from sklearn.model_selection import cross_val_score, StratifiedKFold, KFold
from sklearn.linear_model import LogisticRegression
from sklearn.metrics import accuracy_score
from sklearn.model_selection import GridSearchCV
import statsmodels.formula.api as smf
import textwrap
import scipy.stats as stats
```

```{python}
predict, validate = train_test_split(df, test_size=0.2, random_state=42)
```
```{python}
mod = smf.ols(formula='dv ~ i2 + i3 + i4 + i5 + i6 + i7 + i8 + i9 + i10+i11', data=df)

fit = mod.fit()
fit.summary()

```

```{python}
mod_discovery = smf.ols(formula='dv ~ i2 + i3 + i4 + i5 + i6 + i7 + i8 + i9 + i10 + i11', data=predict)

res_discovery = mod_discovery.fit()
res_discovery.summary()

```

```{python, include = F}
mod_validation = smf.ols(formula='dv ~ i2 + i3 + i4 + i5 + i6 + i7 + i8 + i9 + i10 + i11', data=validate)

res_validation = mod_validation.fit()
```

```{python}
names = ['How likely do you think you are to catch the virus','How badly do you think you will be affected economically if you specifically catch the virus? (for example through loss of work)','How likely do you think it is that a loved one will become infected?','How likely do you think the average person in your state is to become infected?','If you do contract the virus, how likely do you think it is that you will pass it on to someone else','How badly do you think your health will be affected if you do catch the virus?','How badly do you think you will be affected by the global effects of the virus (for example economic recession, reduced healthcare capacity)?','How likely do you think the average person in your neighbourhood is to become infected?','How likely do you think the average person in your country is to become infected?','If you do contract the virus and pass it on to someone else, how badly do you think they would be affected?']


# font type & size
plt.rcParams.update({'font.size': 12})
plt.rcParams["font.family"] = "Times New Roman"

fig, ax = plt.subplots(figsize=(10,14))
fig.set_tight_layout(True)

# DISCOVERY
ax.axvline(0, linestyle=':', color='gray', linewidth=2)

# Points and error bars
ax.errorbar(y=np.arange(len(res_discovery.params[1:])) - 0.1, x=res_discovery.params[1:], 
             xerr=np.abs(res_discovery.conf_int().values[1:, :] - res_discovery.params[1:, np.newaxis]).T, 
             color='black', linewidth=2, fmt='^', label='estimation set')

ax.errorbar(y=np.arange(len(res_validation.params[1:])) + 0.1, x=res_validation.params[1:], 
             xerr=np.abs(res_validation.conf_int().values[1:, :] - res_validation.params[1:, np.newaxis]).T, 
             color='gray', linewidth=2, fmt='s', label='validation set')

# Labels
ax.set_yticks(range(len(names)))
ax.set_yticklabels(['\n'.join(textwrap.wrap(q, 60, break_long_words=False)) for q in names])

## Titles 
ax.set_xlabel('Regression coefficient (+/- 95% CI)')
ax.legend(loc=7)

plt.show()

##save
from PIL import Image
from io import BytesIO
# (1) save the image in memory in tiff format
png1 = BytesIO()
fig.savefig(png1, format= 'png')

# (2) load this image into PIL
png2 = Image.open(png1)

# (3) save as TIFF
png2.save('glm_Daniela_graph3.tiff')
png1.close()
```
